---
title: "Homework 3"
author: Zhongqi Yue
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1
```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and ... columns.

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?

```{r}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```


Let's make a plot

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Let's make a table!!

```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```

Apples vs ice cream..

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```

## Problem 2

Load, tidy, and otherwise wrangle the data.

```{r}
accel_df= (
  read.csv("./data/accel_data.csv")%>% 
  janitor::clean_names() %>% 
    pivot_longer(
      activity_1:activity_1440,
      names_to = "minute_of_a_day",
      names_prefix = "activity",
      values_to = "activity_counts") %>%
  mutate(minute_of_a_day = (as.numeric(as.factor(minute_of_a_day)))) %>% 
    mutate(
      weekday_vs_weekend = case_when(
        day != c("Saturday","Sunday") ~ "weekday",
        day == c("Saturday", "Sunday") ~ "weekend",
        TRUE                           ~ ""))
      
    )
```

After tidy the dataset, there are `r nrow(accel_df)` observations and `r ncol(accel_df)` variables, which are `r names(accel_df)`.

```{r}
accel_df= (
  read.csv("./data/accel_data.csv")%>% 
  janitor::clean_names() %>% 
    pivot_longer(
      activity_1:activity_1440,
      names_to = "minute_of_a_day",
      names_prefix = "activity",
      values_to = "activity_counts") %>% 
  mutate(minute_of_a_day = (as.numeric(as.factor(minute_of_a_day))))%>% 
    mutate(
      weekday_vs_weekend = case_when(
        day != c("Saturday","Sunday") ~ "weekday",
        day == c("Saturday", "Sunday") ~ "weekend",
        TRUE                           ~ "")) %>% 
    group_by(week, day) %>% 
    summarize(total_activity = sum(activity_counts, na.rm = TRUE)) %>%
    mutate(activity_rank = min_rank(total_activity)) %>% 
    	knitr::kable()
      )
```

There is no obvious trend over total activity for each day during these three weeks. However, to some degree, the total activity counts for Tuesday and Wednesday are relatively stable, which are always within 350000~470000 counts across these three weeks. Besides, the rank of the total activity also shows that the rank of Tuesday is relative stable compared to other days, which always ranks 2,3 or 4 within a week. 

```{r}
accel_plot = 
  accel_df= (
  read.csv("./data/accel_data.csv")%>% 
  janitor::clean_names() %>% 
    pivot_longer(
      activity_1:activity_1440,
      names_to = "minute_of_a_day",
      names_prefix = "activity",
      values_to = "activity_counts")
  ) %>% 
  mutate(minute_of_a_day = (as.numeric(as.factor(minute_of_a_day))))

accel_plot %>% 
  	ggplot(aes(x = minute_of_a_day, y = activity_counts, color = day)) + 
	geom_point(alpha = .5)+
  geom_line()
```

The scatter plot indicates the pattern that the male has relatively higher activity counts during the 250~280 minutes for mostly Monday, Friday and Saturday; 550~1125 minutes for mostly Thursday, Friday and Sunday over weeks. For Tuesday and Wednesday, the activity counts are relatively stable over the whole day compared to other days. 

## Problem 3

```{r}
library(p8105.datasets)
library(ggpubr)
data("ny_noaa")
```

This dataset contains `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. The observations are some weather data -- Precipitation, Snowfall,Snow depth, Maximum temperature, Minimum temperature of different weather stations. There are 7 variables, which contains character variables -- id,tmax,tmin; date variable -- date;integer varibales -- prcp, snow, snwd. The missing data in each variable could be an issue since we might not know the weather changes as time went by. 

```{r}
ny_noaa_df=
  ny_noaa %>% 
  separate(date, c("year","month", "day")) %>% 
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    day = as.integer(day)) %>% 
  mutate(tmax = as.numeric(tmax),
         tmin = as.numeric(tmin),
         year = as.character(year),
         month = as.character(month))

ny_noaa_df %>% 
  count(snow) %>% 
  arrange(desc(n))
```

For Snowfall, 0 is the most commonly observed value since it shows up most often among all observed values. Similarly, 25 is the second most commonly observed values; 13 is the third most commonly observed values. 

```{r}
plot_df=
  ny_noaa_df %>% 
  filter(month == c("1","7")) %>%
  group_by(id, year, month, tmax) %>% 
  summarize(
    mean_tmax = mean(tmax)
  )%>% 
  ggplot(aes(x = year, y = mean_tmax, color = id)) + 
	geom_point(alpha = .5)+
  geom_smooth()+
  facet_grid(. ~ month)
```

```{r}
plot1 = 
ny_noaa_df %>% 
  ggplot(aes(x = tmin, y = tmax))+
  geom_bin2d()

plot2 = 
  ny_noaa_df %>% 
  filter(snow > 0,
         snow <100) %>% 
  ggplot(aes(x = year, y = snow, fill = year))+
  geom_violin()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

ggarrange(plot1, plot2, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)
```

